### 2018-09-05 paper/finitecomplete-failures.tex
Small updates

For a given reference process $P$ and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%
\begin{eqnarray}
U_F(p) & = & U_F(p,\varepsilon)
\\
U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p ) \& \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that each test
execution terminates after some $s\in\trc(Q)$ with length $(p+1)$ and
consecutive $\epass$ or $\efail$ event at the latest. The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement as
defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
because $\#s = p$. This is explained further in the next paragraphs.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
empty, so that operand (\ref{eq:ufd}) of the external choice construct
defaults to $\Stop$. In this case, however, $\minaccs(P/s) = \{\varnothing\}$
applies, so the test executes the branch (\ref{eq:ufb}) and terminates with
$\epass$. \fixme{alcc: Oops. This is undefined in CSP.}

>>>>

For a given reference process $P$ and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%
\begin{eqnarray}
U_F(p) & = & U_F(p,\varepsilon)
\\
U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber 
\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events 
$\efail$ or $\epass$ always become visible, if they are the only events
$U_F(p)/s$ is ready to engage in.

The specification of $U_F(p,s)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
If branch (\ref{eq:ufa})
is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
is the last to be entered, the process either accepts another event
$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
and the execution stops without a PASS or FAIL event. Such an execution is also interpreted as FAIL.

The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement as
defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
because $(\#s = p\wedge \minhits(P/s)\neq\varnothing)$. 
This is explained further in the next paragraphs.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
is undefined. Therefore, the associated guard condition contains the requirement
$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken.
Recall that $\minhits(P/s)=\varnothing$ implies  $\minaccs(P/s) = \{\varnothing\}$,
so branches (\ref{eq:ufa}) and (\ref{eq:ufb}) can be taken in this situation.
%%%\fixme{alcc: Oops. This is undefined in CSP.}

### 2018-09-05 paper/introduction.tex
Small updates

Our approach to the definition of CSP fault domains is presented in this
paper. We take advantage of results on model checking of CSP
processes, where the failures semantics of a finite-state CSP process is
represented as a finite normalised transition graph, whose edges are labelled
by the events the process engages in, and whose nodes are labelled by minimal
acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition
graphs do not exceed the order (that is, the number of nodes) of the
reference model's graph by more than a given constant.

>>>>

Our approach to the definition of CSP fault domains is presented in this
paper. We take advantage of results on model checking of CSP
processes, where the failures semantics of a finite-state CSP process is
represented as a finite normalised transition graph, whose edges are labelled
by the events the process engages in, and whose nodes are labelled by minimal
acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:chapter}. The maximal refusals express the
degree of nondeterminism present in a process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition
graphs do not exceed the order (that is, the number of nodes) of the
reference model's graph by more than a given constant.

### 2018-09-05 paper/preliminaries.tex
small changes from the meeting
Small updates

Two forms of choice support branching behaviour. An external choice $P
\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
initial events of $P$ and $Q$. Once synchronisation takes place, the process
that has offered this event is chosen and the other is discarded. In an
internal choice $P \intchoice Q$, the environment does not have an
opportunity to interfere:~the choice is made by the process.
%
\begin{example}\label{example:CSP}
  We consider $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal  choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$. By convention, , if the index set $I$ is
empty, external choices evaluate to the process $\Stop$, which
deadlocks:~does not engage into any event and does not terminate. An iterated
internal choice is not defined for an empty index set. \fixme{alcc: Note the
difference to internal choice. I think this is not a problem here.}

@@

As shown in~\cite{Roscoe:1994:CME:197600}, the failures semantics of any
finite-state CSP process $P$ can be represented by a \emph{normalised
transition graph} $G(P)$ defined by a tuple
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition, that is, an outgoing edge, from $n$ with
label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are
reachable by sequences of edges labelled by $e_1\dots e_k$ and connecting
states $\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
By construction, $s\in\Sigma^*$ is a trace of $P$, if, and only if, there is
a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
$s$. In analogy to $\trc(P)$, we use the notation $\trc(G(P))$ for the set of
finite, initialised paths through $G(P)$, each path represented by its finite
sequence of edge labels. We note that $\trc(P) = \trc(G(P))$. Since $G(P)$ is
normalised, there is a unique node reached by applying the events from $s$
one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well defined.
By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
The graph construction from~\cite{Roscoe:1994:CME:197600} used to define
$G(P)$ for any process $P$ guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
traces $s$ of $P$.

@@

%To see that this approach works only for finite CSP processes, we consider
%the example where $\Sigma$ is infinite. In this case,
%$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
%calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
%defined above. As with refusals, we also use the transition graph-oriented
%notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
%associated with graph state $n$: if $n$ is the state reached in the
%transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
%= \maxrefs(P/s)$.

%Well-formed normalised transition graphs must not refuse an event of the
%initials of a state in {\it every} refusal applicable in this state; more
%formally,
%%
%\begin{equation}
%\label{eq:wellformedg}
%\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
%\exists R\in r(n): e\not\in R
%\end{equation}
%%
%By construction, normalised transition graphs reflect the \emph{failures
%semantics} of finite-state CSP processes:~the traces $s$ of a process are
%defined by the sequences of transitions associated with paths through its
%graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
%r(G(P)/s)$ represent the failures of $P$.

>>>>

Two forms of choice support branching behaviour. An external choice $P
\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
initial events of $P$ and $Q$. Once synchronisation takes place, the process
that has offered this event is chosen and the other is discarded. In an
internal choice $P \intchoice Q$, the environment does not have an
opportunity to interfere:~the choice is made by the process.
%
\begin{example}\label{example:CSP}
  We consider $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal  choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$. By convention, , if the index set $I$ is
empty, external choices evaluate to the process $\Stop$, which
deadlocks:~does not engage into any event and does not terminate. An iterated
internal choice is not defined for an empty index set.

@@

As shown in~\cite{Roscoe:1994:chapter}, the failures semantics of any
finite-state CSP process $P$ can be represented by a \emph{normalised
transition graph} $G(P)$ defined by a tuple
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition, that is, an outgoing edge, from $n$ with
label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:chapter} implies that all nodes $n$ in $N$ are
reachable by sequences of edges labelled by $e_1\dots e_k$ and connecting
states $\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
By construction, $s\in\Sigma^*$ is a trace of $P$, if, and only if, there is
a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
$s$. In analogy to $\trc(P)$, we use the notation $\trc(G(P))$ for the set of
finite, initialised paths through $G(P)$, each path represented by its finite
sequence of edge labels. We note that $\trc(P) = \trc(G(P))$. Since $G(P)$ is
normalised, there is a unique node reached by applying the events from $s$
one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well defined.
By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
The graph construction from~\cite{Roscoe:1994:chapter} used to define
$G(P)$ for any process $P$ guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
traces $s$ of $P$.

@@

% Commented out just to shorten the report.
%To see that this approach works only for finite CSP processes, we consider
%the example where $\Sigma$ is infinite. In this case,
%$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
%calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
%defined above. As with refusals, we also use the transition graph-oriented
%notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
%associated with graph state $n$: if $n$ is the state reached in the
%transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
%= \maxrefs(P/s)$.

% Commented out just to shorten the report.
%Well-formed normalised transition graphs must not refuse an event of the
%initials of a state in {\it every} refusal applicable in this state; more
%formally,
%%
%\begin{equation}
%\label{eq:wellformedg}
%\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
%\exists R\in r(n): e\not\in R
%\end{equation}
%%
%By construction, normalised transition graphs reflect the \emph{failures
%semantics} of finite-state CSP processes:~the traces $s$ of a process are
%defined by the sequences of transitions associated with paths through its
%graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
%r(G(P)/s)$ represent the failures of $P$.
